Experiment 5: Calibration Analysis

### 2024-11-03 group-c/E3-analysis-replication.Rmd

During day 1 of the experiment, participants viewed pairs of images, one of which was always familiar and the other unfamiliar. We calculated a looking score for each participant, defined as the proportion of gaze samples in the ROI of the unfamiliar image out of all the gaze samples that were in either ROI. Gaze samples that were not in either ROI were not included in this analysis. A looking score of 0.5 indicates that participants looked equally often at the familiar and unfamiliar images, while a looking score above 0.5 indicates a preference for the unfamiliar object and a looking score below 0.5 indicate a preference for the familiar object.

Of the `r looking.score.data %>% nrow()` trials in the experiment, `r looking.score.data %>% filter(is.nan(looking.score)) %>% nrow()` had no fixations in either ROI, and so the looking score was unknown. We removed these trials from this analysis.

The mean looking score was `r mean(t.test.data1$M.looking.score)` (*SD* = `r sd(t.test.data1$M.looking.score)`). This was significantly greater than 0.5 (*t*(`r t.test.day1$parameter`) = `r t.test.day1$statistic`, *p* = `r t.test.day1$p.value`), indicating that participants did show a preference for looking at the novel objects.

@@

```{r plot-cumulative-looking-score, echo=FALSE, eval=FALSE, fig.cap="Cumulative looking score over the 5 second exposure during part 2 of day 1. Error bars represent +/- 1 SEM."}
#This (the fig and the DV) isn't referred to in the body of the text, so I'm cutting this for now
ggplot(day.1.looking.by.time.summary, aes(x=t, y=M, ymin=M-SE, ymax=M+SE))+
  geom_pointrange()+
  labs(x = "Time (ms)", y = "Cumulative Looking Score")+
  theme(panel.grid.minor = element_blank())
```

@@

In all of these analyses, we excluded the `r day2.test.data %>% filter(rt_recognition > 10000) %>% nrow()` (out of `r day2.test.data %>% nrow()`) trials where the response time for the recognition judgment was greater than 10 seconds.

Participants correctly identified whether the image was familiar or unfamiliar `r mean(day2.accuracy$accuracy)*100`% (*SD* = `r sd(day2.accuracy$accuracy*100)`) of the time. After excluding the `r length(subjects.100.percent.accurate)` participants who responded correctly to all images, the average confidence rating for correct responses (M = `r confidence.test.data.filtered %>% filter(correct_recognition == TRUE) %>% pull(M.confidence) %>% mean()`; SD = `r confidence.test.data.filtered %>% filter(correct_recognition == TRUE) %>% pull(M.confidence) %>% sd()`) was significantly higher than their average confidence ratings for incorrect responses (M = `r confidence.test.data.filtered %>% filter(correct_recognition == FALSE) %>% pull(M.confidence) %>% mean()`; SD = `r confidence.test.data.filtered %>% filter(correct_recognition == FALSE) %>% pull(M.confidence) %>% sd()`), t(`r confidence.t.test$parameter`) = `r confidence.t.test$statistic`, p = `r confidence.t.test$p.value` . Among the same subset of participants, response times for correct responses (M = `r rt.test.data.filtered %>% filter(correct_recognition == TRUE) %>% pull(M.rt) %>% mean()`, SD = `r rt.test.data.filtered %>% filter(correct_recognition == TRUE) %>% pull(M.rt) %>% sd()`) were also significantly faster than for incorrect responses (M = `r rt.test.data.filtered %>% filter(correct_recognition == FALSE) %>% pull(M.rt) %>% mean()`, SD = `r rt.test.data.filtered %>% filter(correct_recognition == FALSE) %>% pull(M.rt) %>% sd()`), t(`r rt.t.test$parameter`) = `r rt.t.test$statistic` , p = `r rt.t.test$p.value`).

@@

To see whether preferentially looking an the unfamiliar object on day 1 was correlated with confidence and response time for correct responses on Day 2, we computed the correlation coefficient between Day 1 looking scores and Day 2 confidence/RT for each participant. Following the original analysis, we transformed these values using the Fisher p-to-z transformation. Using one-sample t-tests, we found no significant difference from 0 for the correlation between looking score and confidence ratings, t(`r looking.confidence.correlation.t.test$parameter`) = `r looking.confidence.correlation.t.test$statistic`, p = `r looking.confidence.correlation.t.test$p.value` (excluding the subjects who gave the same confidence judgment for all images), nor the correlation between looking score and RT, t(`r  looking.rt.correlation.t.test$parameter`) = `r looking.rt.correlation.t.test$statistic`, p = `r looking.rt.correlation.t.test$p.value`.

>>>>

We calculated the proportion of gaze samples in the ROI of the unfamiliar image out of all the gaze samples that were in either ROI. Of the `r looking.score.data %>% nrow()` trials in the experiment, `r looking.score.data %>% filter(is.nan(looking.score)) %>% nrow()` had no fixations in either ROI and so were excluded from this analysis.

The mean proportion looks to the novel object was `r mean(t.test.data1$M.looking.score)` (*SD* = `r sd(t.test.data1$M.looking.score)`). This was significantly greater than 0.5 (*t*(`r t.test.day1$parameter`) = `r t.test.day1$statistic`, *p* = `r t.test.day1$p.value`), replicating the finding that participants show a preference for looking at the novel objects.

@@

```{r plot-cumulative-looking-score, echo=FALSE, eval=FALSE, fig.cap="Cumulative looking to novel over the 5 second exposure during part 2 of day 1. Error bars represent +/- 1 SEM.", out.width="50%", fig.align="center"}
#This (the fig and the DV) isn't referred to in the body of the text, so I'm cutting this for now
ggplot(day.1.looking.by.time.summary, aes(x=t, y=M, ymin=M-SE, ymax=M+SE))+
  geom_pointrange()+
  labs(x = "Time (ms)", y = "Cumulative Looking to Novel")+
  theme(panel.grid.minor = element_blank())
```

@@

In Day 2 analyses, we excluded the `r day2.test.data %>% filter(rt_recognition > 10000) %>% nrow()` (out of `r day2.test.data %>% nrow()`) trials where the response time for the recognition judgment was greater than 10 seconds.

Participants correctly identified whether the image was familiar or unfamiliar `r mean(day2.accuracy$accuracy)*100`% (*SD* = `r sd(day2.accuracy$accuracy*100)`) of the time. After excluding the `r length(subjects.100.percent.accurate)` participants who responded correctly to all images, the average confidence rating for correct responses (M = `r confidence.test.data.filtered %>% filter(correct_recognition == TRUE) %>% pull(M.confidence) %>% mean()`; SD = `r confidence.test.data.filtered %>% filter(correct_recognition == TRUE) %>% pull(M.confidence) %>% sd()`) was significantly higher than their average confidence ratings for incorrect responses (M = `r confidence.test.data.filtered %>% filter(correct_recognition == FALSE) %>% pull(M.confidence) %>% mean()`; SD = `r confidence.test.data.filtered %>% filter(correct_recognition == FALSE) %>% pull(M.confidence) %>% sd()`), t(`r confidence.t.test$parameter`) = `r confidence.t.test$statistic`, p = `r confidence.t.test$p.value` . Among the same subset of participants, response times for correct responses (M = `r rt.test.data.filtered %>% filter(correct_recognition == TRUE) %>% pull(M.rt) %>% mean()`, SD = `r rt.test.data.filtered %>% filter(correct_recognition == TRUE) %>% pull(M.rt) %>% sd()`) were also significantly faster than for incorrect responses (M = `r rt.test.data.filtered %>% filter(correct_recognition == FALSE) %>% pull(M.rt) %>% mean()`, SD = `r rt.test.data.filtered %>% filter(correct_recognition == FALSE) %>% pull(M.rt) %>% sd()`), t(`r rt.t.test$parameter`) = `r rt.t.test$statistic` , p = `r rt.t.test$p.value`). These findings replicate the original.

@@

To see whether preferentially looking an the unfamiliar object on day 1 was correlated with confidence and response time for correct responses on Day 2, we computed the correlation coefficient between Day 1 proportion looks to novel and Day 2 confidence/RT for each participant. Following the original analysis, we transformed these values using the Fisher p-to-z transformation. Using one-sample t-tests, failed to replicate the a significant difference from 0 for the correlation between proportion looks to the novel object and confidence ratings, t(`r looking.confidence.correlation.t.test$parameter`) = `r looking.confidence.correlation.t.test$statistic`, p = `r looking.confidence.correlation.t.test$p.value` (excluding the subjects who gave the same confidence judgment for all images), nor for the correlation with RT, t(`r  looking.rt.correlation.t.test$parameter`) = `r looking.rt.correlation.t.test$statistic`, p = `r looking.rt.correlation.t.test$p.value`.

### 2024-11-03 group-c/E3-analysis-roi.Rmd

Above we excluded looks that fell outside the ROIs. In contrast, the original experiment simply coded looks as being to the left or right. This may be more appropriate to the limited resolution of camera-based eye-gaze detection. We re-ran analyses based on this split-half coding criterion. The correlation between looking scores using the ROI method and the halves method is `r ls.cor` (see Figure\ \@ref(fig:E3-roi-correlation-of-looking-score)).

```{r E3-roi-correlation-of-looking-score, echo=FALSE, fig.cap="Correlation between looking scores calculated using ROIs and using screen halves.", out.width="50%", fig.align="center"}
ggplot(looking.score.data %>% filter(!is.nan(looking.score.roi)), aes(x=looking.score.roi, y=looking.score.half))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  labs(x="Looking Score (ROIs)", y="Looking Score (Halves)")+
  theme_bw()+
  coord_fixed()+
  theme(panel.grid=element_blank())
```

@@

```{r E3-roi Plot of cumulative looking score, echo=FALSE, eval = FALSE, fig.cap="Cumulative looking score over the 5 second exposure during part 2 of day 1. Error bars represent +/- 1 SEM.", out.width="50%", fig.align="center"}
#This (the fig and the DV) isn't referred to in the body of the text, so I'm cutting this for now
ggplot(day.1.looking.by.time.summary, aes(x=t, y=M, ymin=M-SE, ymax=M+SE))+
  geom_pointrange()+
  labs(x = "Time (ms)", y = "Cumulative Looking Score (Halves)")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
```

@@

Performance on day 2 remained uncorrelated with day 1 looking scores after switching the coding of gaze. We found no significant different from 0 for the correlation between looking score and confidence ratings, t(`r looking.confidence.correlation.t.test$parameter`) = `r looking.confidence.correlation.t.test$statistic`, p = `r looking.confidence.correlation.t.test$p.value` (excluding the subjects who gave the same confidence judgment for all images), nor the the correlation between looking score and RT, t(`r  looking.rt.correlation.t.test$parameter`) = `r looking.rt.correlation.t.test$statistic`, p = `r looking.rt.correlation.t.test$p.value`.

>>>>

Above we excluded looks that fell outside the ROIs. In contrast, the original experiment simply coded looks as being to the left or right. This may be more appropriate to the limited resolution of camera-based eye-gaze detection. We re-ran analyses based on this split-half coding criterion. The correlation between proportion looks to novel using the ROI method and the halves method is `r ls.cor` (see Figure\ \@ref(fig:E3-roi-correlation-of-looking-score)).

```{r E3-roi-correlation-of-looking-score, echo=FALSE, fig.cap="Correlation between proportion looks-to-novel calculated using ROIs and using screen halves.", out.width="50%", fig.align="center"}
ggplot(looking.score.data %>% filter(!is.nan(looking.score.roi)), aes(x=looking.score.roi, y=looking.score.half))+
  geom_point()+
  geom_smooth(method="lm", formula = y~x)+
  labs(x="Looks-to-Novel (ROIs)", y="Looks-to-Novel (Halves)")+
  theme_bw()+
  coord_fixed()+
  theme(panel.grid=element_blank()) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

```

@@

```{r E3-roi Plot of cumulative looking score, echo=FALSE, eval = FALSE, fig.cap="Cumulative looking score over the 5 second exposure during part 2 of day 1. Error bars represent +/- 1 SEM.", out.width="50%", fig.align="center"}
#This (the fig and the DV) isn't referred to in the body of the text, so I'm cutting this for now
ggplot(day.1.looking.by.time.summary, aes(x=t, y=M, ymin=M-SE, ymax=M+SE))+
  geom_pointrange()+
  labs(x = "Time (ms)", y = "Cumulative Looks-to-Novel (Halves)")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())
```

@@

Performance on day 2 remained uncorrelated with day 1 looks-to-novel after switching the coding of gaze. We found no significant different from 0 for the correlation between looks-to-novel and confidence ratings, t(`r looking.confidence.correlation.t.test$parameter`) = `r looking.confidence.correlation.t.test$statistic`, p = `r looking.confidence.correlation.t.test$p.value` (excluding the subjects who gave the same confidence judgment for all images), nor the the correlation between looks-to-novel and RT, t(`r  looking.rt.correlation.t.test$parameter`) = `r looking.rt.correlation.t.test$statistic`, p = `r looking.rt.correlation.t.test$p.value`.

### 2024-11-03 group-d/E4-analysis-ROIs.Rmd

```{r E4-example-subj-looks-ROI, fig.cap = "Example participant's gaze coordinates categorized into ROIs based on a) image bounding boxes and b) screen quadrants. Magenta points indicate looks that were not categorized into an ROI.", out.width="65%", fig.align="center"}
p_box<-ggplot(eyetracking.data.with.roi %>% filter(subject == '5564e577fdf99b6c901a75a6'))+
  geom_point(aes(x = x, y = y, color = as.factor(which_roi)))+
  scale_y_reverse()+
  theme_bw()+
  theme(legend.position = "none")
p_quad<-ggplot(eyetracking.data.with.quadrant.roi %>% filter(subject == '5564e577fdf99b6c901a75a6'))+
  geom_point(aes(x = x, y = y, color = as.factor(which_roi)))+
  scale_y_reverse()+ 
  theme_bw()+
  theme(legend.position = "none")

>>>>

```{r E4-example-subj-looks-ROI, fig.cap = "Example participant's gaze coordinates categorized into ROIs based on a) image bounding boxes and b) screen quadrants. Magenta points indicate looks that were not categorized into an ROI.", out.width="65%", fig.align="center"}
p_box<-ggplot(eyetracking.data.with.roi %>% filter(subject == '5564e577fdf99b6c901a75a6'))+
  geom_point(aes(x = x, y = y, color = as.factor(which_roi)))+
  scale_y_reverse()+
  theme_bw()+
  theme(legend.position = "none")
p_quad<-ggplot(eyetracking.data.with.quadrant.roi %>% filter(subject == '5564e577fdf99b6c901a75a6'))+
  geom_point(aes(x = x, y = y, color = as.factor(which_roi)))+
  scale_y_reverse()+ 
  theme_bw()+
  theme(legend.position = "none") +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

### 2024-11-03 group-e/E5-analysis-calibration.Rmd
Edited ALL expts to have a section under Calibration that re-analyzes data with >50% calibration.

```{r E5cal, fig.show="hold", out.width="45%", echo=FALSE,eval=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:E5cal-caption)"}
ggplot(subject.calibration.score, aes(x=mean_percent_in_roi)) +
  stat_bin(binwidth = 10) +
  labs(x="Calibration Accuracy\n(% of samples within 200px of calibration point)") +
  theme_classic()+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.title.y = element_blank())
```

To determine whether calibration accuracy influenced our key effects, we calculated the proportion of samples during the task in which the participant was fixating the face they ultimately chose. Calibration accuracy significantly correlated with fixations in both the attractiveness condition (r = `r round(scorescorrs.A$estimate, 2)` [`r round(scorescorrs.A$conf.int, 2)[1]`, `r round(scorescorrs.A$conf.int, 2)[2]`], p = `r round(scorescorrs.A$p.value, 2)`) and the roundness condition (r = `r round(scorescorrs.R$estimate, 2)` [`r round(scorescorrs.R$conf.int, 2)[1]`, `r round(scorescorrs.R$conf.int, 2)[2]`], p = `r round(scorescorrs.R$p.value, 2)`). Inspection of Fig. \@ref(fig:E5calcorr) reveals that this correlation is due to a handful of participants with calibration values below 50%.

@@

```{r E5calcorr, fig.show="hold", out.width="80%", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:E5calcorr-caption)"}
ggplot(scores, aes(x=mean_percent_in_roi, y=prop.in.target.subject, group=task, color=task))+
  geom_point() + geom_smooth(method="lm") +
  labs(x="Calibration Accuracy", y="Proportion Fixating Chosen Face") +
  #theme_classic() +
  theme_bw() +
  theme(#panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        #axis.title.y = element_blank()
        )
```

@@

plot.A<-ggplot(rescore[rescore$task=="Attractiveness",], aes(x=time.window, y=prop.in.target)) +
  geom_point(color="red") + 
  geom_line(data=fit.A.dat, aes(x=x, y=y), color="red", lwd=1.25, linetype = "solid") + 
  scale_x_reverse() + theme_classic() + 
  labs(x="Time until decision (ms)", y="Proportion of gaze in target", color="Task", fill="Task") + ylim(.4, 1)

plot.R<-ggplot(rescore[rescore$task=="Roundness",], aes(x=time.window, y=prop.in.target))+
  geom_point(color="turquoise4") + 
  geom_line(data=fit.A.dat, aes(x=x, y=y), color="red", lwd=1.25, linetype = "dashed") + 
  geom_line(data=fit.R.dat, aes(x=x, y=y), color="turquoise4", lwd=1.25, linetype = "solid") + 
  scale_x_reverse() + theme_classic() + 
  labs(x="Time until decision (ms)", y="Proportion of gaze in target", color="Task", fill="Task") + ylim(.4, 1)

Thus, we re-analyzed the data, removing the participants whose calibration accuracy was not greater than 50%. This slightly improved the fits of the sigmoids (Attractiveness: R^2^ = `r round(R2.A,2)`; Roundness: R^2^ = `r round(R2.R,2)`). However, the difference between sigmoids remained non-significant using the Kolmogorov-Smirnov test (D = `r round(KS$statistic, 2)`, p = `r round(KS$p.value, 2)`). Descriptively, the results do not look substantially different (Fig. \@ref(fig:groupEMainrev)).

>>>>

```{r E5cal, fig.show="hold", out.width="45%", echo=FALSE,eval=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:E5cal-caption)", , out.width="50%", fig.align="center"}
ggplot(subject.calibration.score, aes(x=mean_percent_in_roi)) +
  stat_bin(binwidth = 10) +
  labs(x="Calibration Accuracy\n(% of samples within 200px of calibration point)") +
  theme_classic()+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.title.y = element_blank()) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

To determine whether calibration accuracy influenced our key effects, we calculated the proportion of samples during the task in which the participant was fixating the face they ultimately chose. Calibration accuracy significantly correlated with fixations in the attractiveness condition (r = `r round(scorescorrs.A$estimate, 2)` [`r round(scorescorrs.A$conf.int, 2)[1]`, `r round(scorescorrs.A$conf.int, 2)[2]`], p = `r round(scorescorrs.A$p.value, 2)`) but not the roundness condition (r = `r round(scorescorrs.R$estimate, 2)` [`r round(scorescorrs.R$conf.int, 2)[1]`, `r round(scorescorrs.R$conf.int, 2)[2]`], p = `r round(scorescorrs.R$p.value, 2)`). Inspection of Fig. \@ref(fig:E5calcorr) reveals that this correlation is due to a handful of participants with calibration values below 50%.

@@

```{r E5calcorr, fig.show="hold", out.width="80%", echo=FALSE, message=FALSE, warning=FALSE, fig.cap="(ref:E5calcorr-caption)", out.width="50%", fig.align="center"}
ggplot(scores, aes(x=mean_percent_in_roi, y=prop.in.target.subject, group=task, color=task))+
  geom_point() + geom_smooth(method="lm") +
  labs(x="Calibration Accuracy", y="Proportion Fixating Chosen Face") +
  #theme_classic() +
  theme_bw() +
  theme(#panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        #axis.title.y = element_blank()
        ) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )
```

@@

plot.A<-ggplot(rescore[rescore$task=="Attractiveness",], aes(x=time.window, y=prop.in.target)) +
  geom_point(color="red") + 
  geom_line(data=fit.A.dat, aes(x=x, y=y), color="red", lwd=1.25, linetype = "solid") + 
  scale_x_reverse() + theme_classic() + 
  labs(x="Time until decision (ms)", y="Proportion of gaze in target", color="Task", fill="Task") + ylim(.4, 1) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

plot.R<-ggplot(rescore[rescore$task=="Roundness",], aes(x=time.window, y=prop.in.target))+
  geom_point(color="turquoise4") + 
  geom_line(data=fit.A.dat, aes(x=x, y=y), color="red", lwd=1.25, linetype = "dashed") + 
  geom_line(data=fit.R.dat, aes(x=x, y=y), color="turquoise4", lwd=1.25, linetype = "solid") + 
  scale_x_reverse() + theme_classic() + 
  labs(x="Time until decision (ms)", y="Proportion of gaze in target", color="Task", fill="Task") + ylim(.4, 1) +     theme(
    panel.grid.minor = element_blank(),
    axis.title = element_text(size=14),      # Axis titles
    axis.text = element_text(size=14),       # Axis labels
    legend.title = element_text(size=14),    # Legend title
    legend.text = element_text(size=14)      # Legend text
  )

```
#### Re-analysis After Exclusions
As in the previous experiments, we re-analyzed the data, removing the participants whose calibration accuracy was not greater than 50%. This slightly improved the fits of the sigmoids (Attractiveness: R^2^ = `r round(R2.A,2)`; Roundness: R^2^ = `r round(R2.R,2)`). However, the difference between sigmoids remained non-significant using the Kolmogorov-Smirnov test (D = `r round(KS$statistic, 2)`, p = `r round(KS$p.value, 2)`). Descriptively, the results do not look substantially different (Fig. \@ref(fig:groupEMainrev)).
